### 2022-01-19 plos_latex_template.tex
edits to Rules 8–10 ~ @jananiravi

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Unfortunately, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some of them have to deal with several of these on a daily basis.
% While no isolated initiative can drastically alter xx, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles. 
Therefore, recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, is paramount to inclusion, and will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
% These are the groups that warrant deliberate efforts that ensure fair inclusion.

As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g. explicit discrimination, participation fees).
This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
Express this vision for your conference in a diversity statement. 
Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (e.g. a gender distribution of your speakers that is representative of the general population, the presence of racialized people (e.g. Black people) among the organizers, speakers, and attendees, or participation from key geographic regions). 
These numbers should guide you and help achieve your vision.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
Virtual conferences are more inclusive, removing barriers like costs of registration, transport, accommodation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}. 
After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to retain an online component in conferences, either hosting conferences completely virtual or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
It is important to note that while being more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
Key decisions must be made early on during the planning phase \cite{levitisCenteringInclusivityDesign2021}. For instance, the chosen conference platforms (e.g. for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen reader users, geopolitical restrictions, and bandwidth requirements. 
\emoji{bulb} The organizers should also decide how to incorporate geographic regions within the schedule of the conference by asking for pre-recorded videos, recording sessions, and sharing them during the conference. Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly to different time zones; or having time-zone hubs in each day and repeating the events in each (live for one/few and recorded versions for others). 
A flexible schedule and recorded material will allow many more attendees to partake in the conference, e.g. people with schedules conflicting with the live conference, those who cannot take off, or others who are caretakers.

@@

If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format /cite{bajpai_towards_2021}.

@@

\subsection*{Rule 8: Express the welcoming spirit in your communication strategy}
% media
\label{rule_communication}

%Rocío: Main idea: Include people in your communication 
The communication strategy of your conference should aim to increase the number and diversity of participants, and be inclusive by design.
Actively reach out and promote the conference to people who have been systematically excluded. 
Publish the diversity statement, the Code of Conduct, the accessibility guidelines, and  options for financial support (see \textbf{Rule 9} about financial resources), to emphasize that everybody is welcome in the event.
Try to come up with creative ways to show that everyone is seen, respected, and welcome. For example, useR! 2021 created a mascot for the conference wearing different scarves reflecting folks from different groups that the conference wanted to reach (Fig. 2). 
The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize in the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).
Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering personalized help to navigate it, or if captions will be available for some talks but not all.

%Rocío: Main idea: Use inclusive language
Inclusive language—language free from words, phrases or tones that reflect prejudiced or discriminatory views of particular people or groups—should be used in all communications \cite{hallDesigningDiversityInclusion2019}. 
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and with the terms that should be avoided. %
Do not expect minoritized people to teach you—it's not their role—, and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
% links, examples, this changes with the context
%practice: create a standards in inclusive language for your community, to build collective knowledge than can be passed to

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR. %RJ: Need to acknowledge Francisco's contribution.}
\label{fig:marmots}
\end{figure}

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, allow the possibility of a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than non-free events \cite{eventbrite_ultimate_2017}. 
%Other costs can be prohibitive-> scholarships
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying because they think they don't stand a chance). 
Conferences usually ask for cover letters or applications to assign these scholarships, and this can be a time-consuming, emotionally-demanding task. 
Simplify as much as you can the process of asking for financial support; 
people who do not have the time (e.g. because they have family responsibilities) might decide not to apply if the process is overtly complicated. 
Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups. Whenever possible, facilitate ways around money transfer (e.g. book flights, hotel reservations, waive conference fees).

@@

A conference is a critical community-building space, and can foster change in future events and the community itself. The organizers of a conference edition and members of selection committees should do their best to pass their knowledge on best practices and problems encountered to future organizers.

@@

% Rocío: main idea: assess how inclusive was the conference
Assess whether equity and inclusion goals were met during the conference. 
You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and others (e.g. countries of origin, or preferred language).
Other indicators may need a special survey asking for additional information and feedback about the conference. 
In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or did not work. 
% Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.

% Pass the information
Document your processes, outcomes, and lessons learned in blog posts, reports, and presentations. 
Make sure to share this information with future organizers, as well as the contact information of the people you worked with—if they agree to it—, to give them a starting point to improve inclusion in the following editions of the conference.
If you are part of a stable meetings committee, encourage organizers to follow these rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria. 
%something about the past - shoulder of people before, federico

>>>>

% Rocío: Main idea: what is diversity and why does it matter? (inequalities)
Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Unfortunately, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia; some of them have to deal with several of these on a daily basis.
% While no isolated initiative can drastically alter xx, building a more diverse and inclusive conference starts by recognizing that these inequalities have systematically excluded whole groups of people from academia and scientific and professional circles. 
Therefore, recognizing this systematic exclusion, particularly in our fields and in our scientific or professional communities, is paramount to inclusion, and will help identify the most marginalized and discriminated subgroups \cite{timperleyHeMoanaPukepuke2020}.
% These are the groups that warrant deliberate efforts that ensure fair inclusion.
As a starting point, the meeting organizers could identify which groups warrant deliberate efforts to ensure fair inclusion, and determine the contributing factors (e.g. explicit discrimination, participation fees).
This analysis should guide the vision of diversity and inclusion for your conference and help design strategies to achieve it. 
Express this vision for your conference in a diversity statement. 
Next, use this vision and the set strategies to define measurable indicators and goals of diversity and inclusion (e.g. a gender distribution of your speakers that is representative of the general population, the presence of racialized people (e.g. Black people) among the organizers, speakers, and attendees, or participation from key geographic regions). 
These numbers should guide you and help achieve your vision.

@@

% Rocío: main idea of the paragraph: online conferences can open opportunities for inclusion
Virtual conferences are more inclusive, removing barriers like costs of registration, transport, accommodation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}. 
After experiencing virtual conferences mostly due to the COVID pandemic, there is a widespread call to retain an online component in conferences, either hosting conferences completely virtual or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
It is important to note that while being more inclusive, an online component does not guarantee inclusion and accessibility for everyone.
Key decisions must be made early on during the planning phase \cite{levitisCenteringInclusivityDesign2021}. For instance, the chosen conference platforms (e.g. for submission, registration, chatting system, live streaming, etc.) must account for accessibility to accommodate screen-reader users, geopolitical restrictions, and bandwidth requirements. 
\emoji{bulb} The organizers should also decide how to incorporate geographic regions within the schedule of the conference by asking for pre-recorded videos, recording sessions, and sharing them during the conference. Some possibilities for scheduling are: having different days dedicated to different time zones; having each day with the same conference hours, and scheduling time slots that are friendly to different time zones; or having time-zone hubs in each day and repeating the events in each (live for one/few and recorded versions for others). 
A flexible schedule and recorded material will allow many more attendees to partake in the conference, e.g. people with schedules conflicting with the live conference, those who cannot take off, or others who are caretakers.

@@

If you are organizing a hybrid event instead, you should decide on the relative importance that the online and in-person components will take. The most inclusive practice would be to give equal weight to both components, articulating them well and maximizing the experience of everyone attending your conference regardless of the preferred format \cite{bajpai_towards_2021}.

@@

\subsection*{Rule 8: Convey the welcoming spirit in your communication strategy}
% media
\label{rule_communication}

%Rocío: Main idea: Include people in your communication 
The communication strategy of your conference should aim to increase the number and diversity of participants, and be inclusive by design.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Publish the diversity statement, Code of Conduct, accessibility guidelines, and options for financial support (see \textbf{Rule 9} about financial resources) to emphasize that everybody is welcome to join the event.
Try to develop creative ways to show that everyone is seen, respected, and welcome. For example, useR! 2021 created a mascot for the conference wearing different scarves reflecting folks from various minoritized groups that the conference wanted to reach (\textbf{Fig. 2}). 
The diverse members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).
Be transparent with potential attendees and communicate limitations of your conference to let them know what to expect, and ways in which you will try to mitigate these issues. For instance, inform if the conference platform is not completely screen-reader friendly, let people know if you are offering personalized help to navigate it, or if captions will be available for some talks but not all.

%Rocío: Main idea: Use inclusive language
\emoji{bulb} Inclusive language — language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups — should be used in all communications \cite{hallDesigningDiversityInclusion2019}. 
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, and other terms that should be avoided. %
Do not expect minoritized people to teach you—it's not their role—, and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms. 
% links, examples, this changes with the context
%practice: create a standards in inclusive language for your community, to build collective knowledge than can be passed to

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing diverse groups that are part of the large R community: (from left to right) MiR (Minorities in R), LGBTQAI+, R-Ladies, transgender people, and AfricaR.} %RJ: Need to acknowledge Francisco's contribution.
\label{fig:marmots}
\end{figure}

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, allow the possibility of a 'pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
%Other costs can be prohibitive-> scholarships
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying because they think they don't stand a chance). 
Conferences usually ask for cover letters or applications to assign these scholarships, which can be time-consuming and emotionally demanding. 
\emoji{bulb} Simplifying and deconstructing the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because they have family responsibilities) might greatly improve the probability of minoritized participants applying for the much-needed support. 
Bear in mind that transferring money internationally is a cumbersome administrative process that can significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees).

@@

A conference is a critical community-building space, and can foster change in future events and the community itself. The organizers of a conference edition and members of selection committees should do their best to pass along their knowledge on best practices and problems encountered to future organizers.

@@

% Rocío: main idea: assess how inclusive was the conference
As conference organizers, one of the key post-conference tasks is to analyze conference attendee, participant, and organizer data including their experiences. It is essential to assess whether equity and inclusion goals were met during the conference, as a result of the steps taken. 
\emoji{bulb} You can rely on the diversity indicators defined at the beginning of the planning process (\textbf{Rule 1}). 
It is likely that you will be able to calculate some of your indicators directly from information that you have already collected from your organizing team, speakers, and participants (e.g. countries of origin, or preferred language).
Other indicators may need a special survey asking for additional information and feedback about the conference. 
In the survey, providing options for open answers will allow people to express their opinions freely, and may give you more insights into how/why your inclusion practices worked or not. 
% Do not expect everything to be perfectly inclusive, or to encompass all the dimensions of diversity at once. %You're part of a process and contributing to structural change.

% Pass the information
%\emoji{bulb} Practical tips
Document your processes, outcomes, and lessons learned through blog posts, reports, and presentations. 
Make sure you share this information with future organizers, as well as the contact information of the people you worked with (if they agree to it), to give them a good well-informed baseline to improve inclusion in the following editions of the conference.
If you are part of a stable meeting committee (overseeing multiple editions), encourage organizers to follow these Rules and set up new standards for inclusion: be explicit on the inclusive spirit of the conference and the practices that future organizers should commit to when publishing the call for conference organization and the selection criteria. 
%something about the past - shoulder of people before, federico
